\textbf{On the Common Logical Structure of Classical and Quantum Mechanics}

### 2021-06-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: addressing Christine's TODO

In order to begin our discussion, let us consider in the first place a set $X$: in the quantum case it will correspond to the set of all possible quantum states; similarly, in the classical case it will correspond to the set of all possible probability distributions over phase space.\footnote{We should stress, however, that $\sigma$-algebras are defined independently of what the elements of $X$ actually are. For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.} By definition, a $\sigma$-algebra is a collection of subsets of $X$ closed under countable union, countable intersection and complement with respect to $X$; the closure condition requires that  performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra. Interestingly, these sets can be understood to represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Namely, if $U \subseteq X$, the statement ``$x$ is in $U$'' tells us that the state $x$ must be one of those delimited by $U$. Then, by virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets. Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement [TODO: For statements, should this be complement or negation?]. Hence, the $\sigma$-algebra generates a Boolean algebra of statements.\footnote{Note that the interpretation used here is exactly the same used in probability theory. In fact, a probability space is defined by three elements: a sample space $\Omega$ which represents the set of all possible outcomes; a $\sigma$-algebra over $\Omega$ which represents all events, all propositions; and a measure that assigns a probability to each event.}

@@

Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''. This is \emph{not} equivalent to the statement ``the particle is in the eigenstate $x_0$'' [TODO: Does it make sense to say that the eigenstate, rather than the eigenvalue, is $x_0$? Is different notation needed to make it a state?]. This latter proposition will be true only if the whole function is at $x_0$. That is, if $x_0$ is the support of the wave function. The former is much less stringent: the wave function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$. This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).

@@

Now that we have argued that quantum mechanics can be given an algebra that looks more like [TODO: Should a stronger phrase than "looks more like" be used?] the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic. Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory. Thus, let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space. As already stated, we can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$. This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$. Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear, we are not arguing that this construction is the most appropriate or useful. Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.

>>>>

In order to begin our discussion, let us consider in the first place a set $X$: in the quantum case it will correspond to the set of all possible quantum states; similarly, in the classical case it will correspond to the set of all possible probability distributions over phase space.\footnote{We should stress, however, that $\sigma$-algebras are defined independently of what the elements of $X$ actually are. For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.} By definition, a $\sigma$-algebra is a collection of subsets of $X$ closed under countable union, countable intersection and complement with respect to $X$; the closure condition requires that  performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra. Interestingly, these sets can be understood to represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Namely, if $U \subseteq X$, the statement ``$x$ is in $U$'' tells us that the state $x$ must be one of those delimited by $U$. Then, by virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets. Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and negation. Hence, the $\sigma$-algebra generates a Boolean algebra of statements.\footnote{Note that the interpretation used here is exactly the same used in probability theory. In fact, a probability space is defined by three elements: a sample space $\Omega$ which represents the set of all possible outcomes; a $\sigma$-algebra over $\Omega$ which represents all events, all propositions; and a measure that assigns a probability to each event.}

@@

Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''. This is \emph{not} equivalent to the statement ``the particle is in the eigenstate of position corresponding to $x_0$'' This latter proposition will be true only if the whole function is at $x_0$. That is, if $x_0$ is the support of the wave function. The former is much less stringent: the wave function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$. This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).

@@

Now that we have argued that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic. Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory. Thus, let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space. As already stated, we can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$. This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$. Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear, we are not arguing that this construction is the most appropriate or useful. Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.
